---
title: "haheho"
author: "Pandu Jatayu Satrio Aji"
date: "2023-12-04"
output: html_document
---


### 1. Load Library Needed
```{r}
library(rjags)
```

### 2. Load the Data
```{r}
data <- read.csv('penguins_lter.csv')

study_name <- data[,1]
sample_num <- data[,2]
species <- data[,3]
region <- data[,4]
island <- data[,5]
stage <- data[,6]
individual_id <- data[,7]
clutch_completion <- data[,8]
date_egg <- data[,9]
culmen_length <- data[,10]
culmen_depth <- data[,11]
flipper_length <- data[,12]
body_mass <- data[,13]
sex <- data[,14]
delta_N <- data[,15]
delta_C <- data[,16]
comment <- data[,17]
```

### 3. Encoding (Convert Categorical Data Type to Numeric that We Need)
```{r}
data$Sex[data$Sex == "MALE"] <- 1
data$Sex[data$Sex == "FEMALE"] <- 0
data$Sex <- as.numeric(data$Sex, coerce=TRUE)
```

```{r}
data$Clutch.Completion[data$Clutch.Completion == "Yes"] <- 1
data$Clutch.Completion[data$Clutch.Completion == "No"] <- 0
data$Clutch.Completion <- as.numeric(data$Clutch.Completion, coerce=TRUE)
```

```{r}
data$Flipper.Length..mm. <- as.numeric(data$Flipper.Length..mm., coerce=TRUE)
data$Body.Mass..g. <- as.numeric(data$Body.Mass..g., coerce=TRUE)
```

```{r}
str(data)
```


```{r}
View(data)
```

```{r}
clutch_completion <- data[,8]
date_egg <- data[,9]
culmen_length <- data[,10]
culmen_depth <- data[,11]
flipper_length <- data[,12]
body_mass <- data[,13]
sex <- data[,14]
delta_N <- data[,15]
delta_C <- data[,16]
```

### 4. Plot Data
```{r}
Y <- culmen_length
X <- cbind(clutch_completion,
           culmen_depth, flipper_length, body_mass,
           sex, delta_N, delta_C)
names <- c("Clutch completion", "Culmen depth",
           "Flipper length", "Body mass", "Sex", "Delta N", "Delta C")

# Remove observations with NA values

junk <- is.na(rowSums(X))
Y <- Y[!junk]
X <- X[!junk,]
study_name <- study_name[!junk]
sample_num <- sample_num[!junk]
species <- species[!junk]
region <- region[!junk]
island <- island[!junk]
stage <- stage[!junk]
individual_id <- individual_id[!junk]
date_egg <- date_egg[!junk]
comment <- comment[!junk]

# Standardize the covariates

X <- as.matrix(scale(X))
```

```{r}
names <- c("Clutch completion", "Culmen depth",
           "Flipper length", "Body mass", "Sex", "Delta N", "Delta C")
for(j in 1:7){
  boxplot(X[,j]~Y, main=names[j])
}
```


### 5.Put Data in JAGS Format
```{r}
n <- length(Y)
p <- ncol(X)

data <- list(Y=Y, X=X, n=n, p=p)
params <- c("beta")

burn <- 10000
n.iter <- 20000
thin <- 10
n.chains <- 3
```

### 6. Fit the Uninformative Gaussian Model (Model 1)
```{r}
model_string <- textConnection("model{
                # Likelihood
                for(i in 1: n){
                  Y[i] ~ dnorm(alpha+inprod(X[i,], beta[]), taue)
                }
                # Priors
                for(j in 1: p){
                  beta[j] ~ dnorm(0, 0.001)
                }
                alpha ~ dnorm(0, 0.001)
                taue ~ dgamma(0.1, 0.1)
                
                # for WAIC calculations
                for(i in 1:n){
                like[i] <- dnorm(Y[i], inprod(X[i,], beta[]), taue)
                }
}")

model <- jags.model(model_string, data = data, n.chains = n.chains, quiet = TRUE)
update(model, burn, progress.bar = "none")
samples1 <- coda.samples(model, variable.names = params, n.thin = thin, n.iter = n.iter)

par(mar=c(2,2,2,2))
plot(samples1)
```


```{r}
sum <- summary(samples1)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### 7. Summarize the marginal distributions (Model 1)
```{r}
beta <- NULL
  for(l in 1:n.chains){
    beta <- rbind(beta,samples1[[l]])
  }
  colnames(beta) <- names
  for(j in 1:7){
    hist(beta[,j],xlab=expression(UninformativeGaussianModel),ylab="Posterior density",
    breaks=100,main=names[j])
}

```

```{r}
library(knitr)
Inc_Prob <- apply(beta!=0,2,mean)
Q <- t(apply(beta,2,quantile,c(0.5,0.05,0.95)))
out <- cbind(Inc_Prob,Q)
kable(round(out,2))
```

### 8. Compute the posterior probability of each model (Model 1)
```{r}
model_a <- "Intercept"
for(j in 1:7){
  model_a <- paste(model_a,ifelse(beta[,j]==0,"","+"))
  model_a <- paste(model_a,ifelse(beta[,j]==0,"",names[j]))
}
model_a[1:5]
```

```{r}
beta[1:5,]
```
### 9. SSVS result (this model dominate the posterior probability)
```{r}
model_probs <- table(model_a)/length(model_a)
model_probs <- sort(model_probs,dec=T)
round(model_probs,3)
```


### 10. Convergence diagnostics model 1
```{r}
autocorr.plot(samples1)
```

#### Numerical diagnostics
```{r}
# Low autocorrelation indicates convergence
autocorr(samples1[[1]],lag=1)
```

```{r}
# ESS over 1000 indicates convergence
effectiveSize(samples1)
```

```{r}
# Gelman-Rubin statistic
# R less than 1.1 indicates convergence
gelman.diag(samples1)
```

```{r}
# Geweke diagnostic
# |z| less than 2 indicates convergence
geweke.diag(samples1[[1]])
```
### 11. Model Performance (Model 1)
```{r}
# DIC
dic1 <- dic.samples(model, n.iter = 20000, n.thin= 10, progress.bar = "none")
dic1
```

```{r}
# WAIC
waic1 <- coda.samples(model, variable.names=c("like"), n.iter = n.iter, progress.bar="none")
beta1 <- waic1[[1]]
fbar1 <- colMeans(beta1)
P1 <- sum(apply(log(beta1), 2, var))
waic_1 <- -2*sum(log(fbar1)) + 2*P1
waic_1
```

### 12. Plot Data (Model 2 with only high convergence)
```{r}
Y2 <- culmen_length
X2 <- cbind(clutch_completion,
           culmen_depth,
           delta_N, delta_C)
names2 <- c("Clutch completion", "Culmen depth",
            "Delta N", "Delta C")

# Remove observations with NA values

junk <- is.na(rowSums(X2))
Y2 <- Y2[!junk]
X2 <- X2[!junk,]
study_name <- study_name[!junk]
sample_num <- sample_num[!junk]
species <- species[!junk]
region <- region[!junk]
island <- island[!junk]
stage <- stage[!junk]
individual_id <- individual_id[!junk]
date_egg <- date_egg[!junk]
comment <- comment[!junk]

# Standardize the covariates

X2 <- as.matrix(scale(X2))
```

### 13. Put Data in JAGS Format (model 2)
```{r}
n2 <- length(Y2)
p2 <- ncol(X2)

data2 <- list(Y=Y2, X=X2, n=n2, p=p2)
params <- c("beta")

burn <- 10000
n.iter <- 20000
thin <- 10
n.chains <- 3
```

### 14. Fit the Uninformative Gaussian Model (model 2)
```{r}
model_string <- textConnection("model{
                # Likelihood
                for(i in 1: n){
                  Y[i] ~ dnorm(alpha+inprod(X[i,], beta[]), taue)
                }
                # Priors
                for(j in 1: p){
                  beta[j] ~ dnorm(0, 0.001)
                }
                alpha ~ dnorm(0, 0.001)
                taue ~ dgamma(0.1, 0.1)
                
                # For compute WAIC
                for(i in 1:n){
                like[i] <- dnorm(Y[i], inprod(X[i,], beta[]), taue)
                }
                
}")

model2 <- jags.model(model_string, data = data2, n.chains = n.chains, quiet = TRUE)
update(model2, burn, progress.bar = "none")
samples2 <- coda.samples(model2, variable.names = params, n.thin = thin, n.iter = n.iter)

par(mar=c(2, 2, 2, 2))
plot(samples2)
```

```{r}
sum <- summary(samples2)
rownames(sum$statistics) <- names2
rownames(sum$quantiles) <- names2
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### 15. Summarize the marginal distributions (model 2)
```{r}
beta <- NULL
  for(l in 1:n.chains){
    beta <- rbind(beta,samples2[[l]])
  }
  colnames(beta) <- names2
  for(j in 1:4){
    hist(beta[,j],xlab=expression(UninformativeGaussianModel2),ylab="Posterior density",
    breaks=100,main=names2[j])
}

```

```{r}
library(knitr)
Inc_Prob <- apply(beta!=0,2,mean)
Q <- t(apply(beta,2,quantile,c(0.5,0.05,0.95)))
out <- cbind(Inc_Prob,Q)
kable(round(out,2))
```

### 16. Compute the posterior probability of each model
```{r}
model_b <- "Intercept"
for(j in 1:4){
  model_b <- paste(model_b,ifelse(beta[,j]==0,"","+"))
  model_b <- paste(model_b,ifelse(beta[,j]==0,"",names2[j]))
}
model_b[1:5]
```

```{r}
beta[1:5,]
```

```{r}
model_probs <- table(model_b)/length(model_b)
model_probs <- sort(model_probs,dec=T)
round(model_probs,3)
```

### 17. Convergence diagnostics model 2
```{r}
autocorr.plot(samples2)
```

#### Numerical diagnostics
```{r}
# Low autocorrelation indicates convergence
autocorr(samples2[[1]],lag=1)
```

```{r}
# ESS over 1000 indicates convergence
effectiveSize(samples2)
```

```{r}
# Gelman-Rubin statistic
# R less than 1.1 indicates convergence
gelman.diag(samples2)
```

```{r}
# Geweke diagnostic
# |z| less than 2 indicates convergence
geweke.diag(samples2[[1]])
```

### 18. Model Performance (Model 2)
```{r}
# DIC
dic2 <- dic.samples(model2, n.iter = 20000, n.thin= 10, progress.bar = "none")
dic2
```

```{r}
# WAIC
waic2 <- coda.samples(model2, variable.names=c("like"), n.iter = n.iter, progress.bar="none")
beta2 <- waic2[[1]]
fbar2 <- colMeans(beta2)
P2 <- sum(apply(log(beta2), 2, var))
waic_2 <- -2*sum(log(fbar2)) + 2*P2
waic_2
```


### 19. Fit the Gaussian Shrinkage Model (additional model)

```{r}
model_string <- textConnection("model{
                # Likelihood
                  for(i in 1:n){
                    Y[i] ~ dnorm(alpha+inprod(X[i,],beta[]),taue)
                  }
                # Priors
                  for(j in 1:p){
                    beta[j] ~ dnorm(0,taue*taub)
                  }
                  alpha ~ dnorm(0,0.001)
                  taue ~ dgamma(0.1, 0.1)
                  taub ~ dgamma(0.1, 0.1)
                  
                  # For compute WAIC
                  for(i in 1:n){
                  like[i] <- dnorm(Y[i], inprod(X[i,], beta[]), taue)
                  }
                }")

model3 <- jags.model(model_string,data = data, n.chains=n.chains,quiet=TRUE)
update(model3, burn, progress.bar="none")
samples3 <- coda.samples(model3, variable.names=params, n.thin=thin, n.iter=n.iter)

par(mar=c(2,2,2,2))
plot(samples3)
```


```{r}
names <- c("Clutch completion", "Culmen depth",
           "Flipper length", "Body mass", "Sex", "Delta N", "Delta C")
sum <- summary(samples3)
rownames(sum$statistics) <- names
rownames(sum$quantiles) <- names
sum$statistics <- round(sum$statistics,3)
sum$quantiles <- round(sum$quantiles,3)
sum
```

### 20. Convergence diagnostics model 3
```{r}
autocorr.plot(samples3)
```

#### Numerical diagnostics
```{r}
# Low autocorrelation indicates convergence
autocorr(samples3[[1]],lag=1)
```

```{r}
# ESS over 1000 indicates convergence
effectiveSize(samples3)
```

```{r}
# Gelman-Rubin statistic
# R less than 1.1 indicates convergence
gelman.diag(samples3)
```

```{r}
# Geweke diagnostic
# |z| less than 2 indicates convergence
geweke.diag(samples3[[1]])
```

### 21. Model Performance for Additional Model
```{r}
dic3 <- dic.samples(model3, n.iter = 20000, n.thin= 10, progress.bar = "none")
dic3
```

```{r}
waic3 <- coda.samples(model3, variable.names=c("like"), n.iter = n.iter, progress.bar="none")
beta3 <- waic3[[1]]
fbar3 <- colMeans(beta3)
P3 <- sum(apply(log(beta3), 2, var))
waic_3 <- -2*sum(log(fbar3)) + 2*P3
waic_3
```


### Thank you